Activity Monitoring Analysys
============================

# Note

Data file must be in a working directory.

# Loading and preprocessing the data.

```{r}
par(mfrow=c(1, 1))

data <- read.csv("activity.csv", stringsAsFactors = F)
data$DT <- strptime(data$date, format="%Y-%m-%d")
data$WD <- weekdays(data$DT, abbreviate=T)
dx <- data[!is.na(data$steps),]
```

# Mean of the total number of steps taken per day

```{r fig.width=8, fig.height=6}
dx <- data[!is.na(data$steps),]
len <- sum(dx$steps)
d0 <- character(length = len)
k <- 1L
for (i in seq(1, nrow(dx))){
    cc <- dx[i,]
    if (cc$steps > 0) {
        for (j in seq(1, cc$steps)) {
            d0[k] <- cc$date
            k <- k + 1L
        }
    }
}
d0 <- strptime(d0, format="%Y-%m-%d")
hist(d0, breaks="days", format="%Y-%m-%d", col = "red", xlab="Date", ylab="Number of Steps", 
     main="Number of Steps per Day", freq = TRUE)
```

## Mean and median steps taken per day

```{r}
dx <- data[!is.na(data$steps),]
d1 <- split(dx, factor(as.Date(dx$date, format="%Y-%m-%d")))
d2 <- data.frame()
for (c in d1) {
    d2 <- rbind(d2, data.frame(Date = c[1, "DT"], Mean = mean(c$steps), Median = median(c$steps)))
}
d2
```

# Average daily activity pattern

```{r fig.width=8, fig.height=6}
dx <- data[!is.na(data$steps),]
d1 <- split(dx, factor(dx$interval))
d3 <- data.frame()
for (c in d1) {
    d3 <- rbind(d3, data.frame(Interval = c[1, "interval"], Mean = mean(c$steps)))
}

plot(d3$Interval, d3$Mean, type ="l", lwd = 3, col="red", main = "Average Steps per 5 Minutes Interval", 
     xlab = "Interval (minutes from the start of the day)", ylab = "Average steps")
```

## Maximum number of steps in interval

```{r}
d3[order(d3$Mean,decreasing=T),][1,]
```

# Imputing missing values

Since some values are missing, impute the values based on mean for a pair interval/day of the week for existing values.

## Missing values count

```{r}
sum(is.na(data))
```

## Total values count

```{r}
nrow(data)
```

## Results of imputing missing values

```{r fig.width=8, fig.height=6}
dx <- data[!is.na(data$steps),]
dx$key <- paste(dx$interval, dx$WD, sep=";")
d6 <- split(dx, factor(dx$key))
d7 <- data.frame()
for (c in d6) {
    d7 <- rbind(d7, data.frame(key = c[1, "key"], mean = mean(c$steps)))
}

dy <- data[is.na(data$steps),]
dy$key <- paste(dy$interval, dy$WD, sep=";")
st <- numeric()
for (i in 1:nrow(dy)) {
    cc <- dy[i,]
    st <- c(st, d7[d7$key == paste(cc$interval, cc$WD, sep=";"), "mean"])
}
dy$steps <- st     

data1 <- rbind(dx, dy)

# Step 6
dx <- data1
len <- sum(dx$steps)
d0 <- character(length = len)
k <- 1L
for (i in seq(1, nrow(dx))){
    cc <- dx[i,]
    if (cc$steps > 0) {
        for (j in seq(1, cc$steps)) {
            d0[k] <- cc$date
            k <- k + 1L
        }
    }
}
d0 <- strptime(d0, format="%Y-%m-%d")
hist(d0, breaks="days", format="%Y-%m-%d", col = "red", xlab="Date", ylab="Number of Steps", main="Number of Steps per Day", freq = TRUE)

```

Actually there are no significant difference between original and modified dataset histogram.

# Differences in activity patterns between weekdays and weekends

```{r fig.width=8, fig.height=12}
par(mfrow=c(2, 1))

dx <- data[!is.na(data$steps),]

d0 <- dx[(dx$WD == "Sun") | (dx$WD == "Sat"), ]

d1 <- split(d0, factor(d0$interval))
d3 <- data.frame()
for (c in d1) {
    d3 <- rbind(d3, data.frame(Interval = c[1, "interval"], Mean = mean(c$steps)))
}
plot(d3$Interval, d3$Mean, type ="l", lwd = 3, col="blue", main = "Weekend", 
     xlab = "Interval (minutes from the start of the day)", ylab = "Number of Steps")

d0 <- dx[!((dx$WD == "Sun") | (dx$WD == "Sat")), ]
d1 <- split(d0, factor(d0$interval))
d3 <- data.frame()
for (c in d1) {
    d3 <- rbind(d3, data.frame(Interval = c[1, "interval"], Mean = mean(c$steps)))
}
plot(d3$Interval, d3$Mean, type ="l", lwd = 3, col="blue", main = "Weekday", 
     xlab = "Interval (minutes from the start of the day)", ylab = "Number of Steps")
```
